**Machine learning identifies unique taxa differentiating proximal and distal

### 2017-05-25 submission/manuscript.Rmd
discussion

To get around the noisiness created by a diverse set of samples, we employed machine learning algorithms to identify microbes specific to each side.

Random Forest models have been used by our group and others to increase the detection sensitivity of CRC tumors. Indeed, our group found that a classification model that used microbiome data in combination with Fecal Immunohistochemical Test (FIT) results could correctly identify both carcinoma and adenoma lesions from communities of stool at exit (Baxter et al 2016) and it performed much better than FIT alone. 
- Could try to predict L vs R tumors from exit stool? long term. maybe. probably not. value of getting data about exit stool comparisons. real issue is diagnosing right though. 
- note that we have FIT for all of these people but they're all below limit of detection (except for 2). when we look at models for CRC, cant predict R vs L, so create models for healthy

Discussion of differentially abundant OTUs, how they fit the ecological regions of the gut (schematic?)
some of the very high OTUs BLAST to stool residents, like that one bacteriodes (OTU20) - not necessarily anything interesting there.

@@

what is rate of colonization with these bugs in healthy in literature?

>>>>

To get around the noisiness created by a diverse set of samples, we employed machine learning algorithms to identify microbes specific to each side. For each comparison we identified top X OTUs that were strongly predictive of one site or another. Generally, OTUs identified in each location were consistent with known physiological gradients along the gut axis (cite nature review). For instance, the right mucosa harbored mucosa-associated facultative anaerobes such as _Actinomyces_ and _Enterobacteraceae_ and the aerobic _Psuedomonas_ consistent with the highest oxygen regions of the colon (cite). The left mucosa was far more likely to host strictly anaerobic species such as _Porphyromonas_, _Anaerococcus_, _Finegoldia_ and _Peptoniphilus_. The model was less effective at classifying the right and left lumenal contents, probably because the samples are arguably composed of the same material.

other factors besides oxygen? pH? 
offer another model/schematic?
some of the very high OTUs BLAST to stool residents, like that one bacteriodes (OTU20) - not necessarily anything interesting there.

what is rate of colonization with these bugs in healthy in literature? how did these bugs factor into our model?

@@

how does bowel prep fit into all of this

Random Forest models have been used by our group and others to increase the detection sensitivity of CRC tumors. Indeed, our group found that a classification model that used microbiome data in combination with Fecal Immunohistochemical Test (FIT) results could correctly identify both carcinoma and adenoma lesions from communities of stool at exit (Baxter et al 2016) and it performed much better than FIT alone. 
- Could try to predict L vs R tumors from exit stool? long term. maybe. probably not. value of getting data about exit stool comparisons. real issue is diagnosing right though. 
- note that we have FIT for all of these people but they're all below limit of detection (except for 2). when we look at models for CRC, cant predict R vs L, so create models for healthy. can also cite niels stuff showing that FIT is representative of stool community. could it also be representative of left? 
- LINK TO EXIT STOOL COMPARISONS.
